---
title: "Tidy Tuesday (2021/10/19)"
output: pdf_document
---

```{R warning=FALSE, message=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(ggmap)
```

# Data Tidying

Loading in the data and specifying specific data types for each column. `pumpkins` will serve as the main `tibble` for the remainder of the analysis.

```{R warning=FALSE, message=FALSE}
set.seed(20211019)

tuesdata = tidytuesdayR::tt_load('2021-10-19')

pumpkins = tuesdata$pumpkins %>%
  select(-variety) %>%
  mutate(
    place = as.integer(place),
    weight_lbs = as.double(weight_lbs),
    grower_name = as.factor(grower_name),
    city = as.factor(city),
    state_prov = as.factor(state_prov),
    country = as.factor(country),
    gpc_site = as.factor(gpc_site),
    seed_mother = as.factor(seed_mother),
    pollinator_father = as.factor(pollinator_father),
    ott = as.double(ott),
    est_weight = as.double(est_weight),
    pct_chart = as.double(pct_chart)
  )
```

\newpage

# Top Seed Mothers

*Which seed mothers are best at producing the highest-placed pumpkins?* Using `seed_mother`s that appear more than 26 times (number chosen solely for maximal plotting information), we notice that most `seed_mother`s follow a similar trend: the higher the weight, the greater the place (observed here as a negative relationship). Growers should look for seeds that provide higher places more consistently than other seeds. Seeds such as `211 MacKinnon`, `302 Kent`, and `316 Edwards` appear to almost guarantee a high position if the pumpkin exceeds a certain weight, whereas other seeds like `2145 McMullen`, `1985 Miller` and `1756 Lancaster` show a linear relationship between weight and placement.

```{r}
top_seed_mothers = pumpkins %>%
  select(place, weight_lbs, city, state_prov, gpc_site, seed_mother, pollinator_father) %>%
  filter((!is.na(seed_mother)) & (seed_mother != 'Unknown') & (seed_mother != 'unknown')) %>%
  group_by(seed_mother) %>%
  summarize(n = n()) %>%
  filter(n > 26) %>%
  arrange(desc(n))

top_seed_mothers

pumpkins %>%
  filter(seed_mother %in% top_seed_mothers$seed_mother) %>%
  ggplot(aes(x=weight_lbs, y=place, colour=seed_mother)) +
    geom_point(position='jitter', show.legend=FALSE) +
    facet_wrap(~ seed_mother)
```

\newpage

# Top Region Producers

*Do specific regions produce better than others?* If up to me, I'd rather have `latitude-longitude` data - which is totally possible, I just couldn't find a convenient solution that provided access to `lat-long` given `city-state` pairings (thanks, Google). However, we notice very similar (almost identical) trends between all regions, implying that competitions are relatively fair across all regions (i.e. one does not gain an advantage competing in one region than they do any other region).

```{R}
new_england = c('Connecticut', 'Maine', 'Massachusetts', 'New Hampshire',
              'Rhode Island', 'Vermont')
mid_atlantic = c('New Jersey', 'New York', 'Pennsylvania')
east_north_central = c('Illinois', 'Indiana', 'Michigan', 'Ohio',
                       'Wisconsin')
west_north_central = c('Iowa', 'Kansas', 'Minnesota', 'Missouri', 'Nebraska',
                       'North Dakota', 'South Dakota')
south_atlantic = c('Delaware', 'Florida', 'Georgia', 'Maryland',
                   'North Carolina', 'South Carolina', 'Virginia',
                   'Washington D.C.', 'West Virginia')
east_south_central = c('Alabama', 'Kentucky', 'Mississippi', 'Tennessee')
west_south_central = c('Arkansas', 'Lousiana', 'Oklahoma', 'Texas')
mountain = c('Arizona', 'Colorado', 'Idaho', 'Montana', 'Nevada', 'New Mexico', 'Utah', 'Wyoming')
pacific = c('Alaska', 'California', 'Hawaii', 'Oregon', 'Washington')

geo_pumps = pumpkins %>%
  filter(country == 'United States') %>%
  select(-c('grower_name', 'country', 'gpc_site', 'seed_mother', 'pollinator_father')) %>%
  mutate(region = ifelse(state_prov %in% new_england, 'New England',
                  ifelse(state_prov %in% mid_atlantic, 'Mid-Atlantic',
                  ifelse(state_prov %in% east_north_central, 'Midwest',
                  ifelse(state_prov %in% west_north_central, 'Midwest',
                  ifelse(state_prov %in% south_atlantic, 'South',
                  ifelse(state_prov %in% east_south_central, 'South',
                  ifelse(state_prov %in% west_south_central, 'South',
                  ifelse(state_prov %in% mountain, 'Pacific',
                  ifelse(state_prov %in% pacific, 'Pacific', 'Undefined')))))))))
         )

geo_pumps %>%
  ggplot() +
  geom_point(aes(x=weight_lbs, y=place, colour=region),
             position='jitter', show.legend=FALSE) + 
  facet_wrap(~ region)
```

\newpage

# Repeat Offenders

*How well do specific groups (based on number of attempts) perform?* It seems like the more attempts one takes, the better off they fair, even if their pumpkins are of less weight. Granted there is increasingly less data supporting this per group, it is worth noting that the relationship between `weight_lbs` and `place` changes per group (where relationships can be clearly observed).

```{r}
repeat_offenders = pumpkins %>%
  group_by(grower_name) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  filter(n >= 5)

pumpkins %>%
  filter(grower_name %in% repeat_offenders$grower_name) %>%
  group_by(grower_name) %>%
  summarize(
    place = mean(place, na.rm=TRUE),
    weight_lbs = mean(weight_lbs, na.rm=TRUE),
    count = n()
  ) %>%
  mutate(attempts = ifelse(count == max(count), 'master',
                    ifelse(max(count) >= count & count > 60, 'elite',
                    ifelse(60 >= count & count > 30, 'pro',
                    ifelse(30 >= count & count > 15, 'amateur',
                    ifelse(15 >= count & count > 5, 'rookie', 'novice')))))) %>%
  ggplot(aes(x=weight_lbs, y=place, colour=attempts)) +
    geom_point() +
    facet_grid(attempts ~ .)
```

